---
title: "Figure 5: Normative development"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/CamRat/CamRat/"
source(paste0(base_dir, "code/03.data_analysis/setup.R"))
figures_dir <- paste0(base_dir, "outputs/figures2.0/Fig5/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/")

## Load data objects figures
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}
```

```{r plot.setup}
## Identify max and min of slope effect size color scale based on maximum absolute slope
degree_slope_scale_max <- df_node_slopes_roi %>% 
  filter(feature == "degree" & term == "age") %>% 
  pull(t_value) %>% 
  abs %>% 
  max %>% 
  round(1)

## Set edge color scale (to be consistent for Figs 3h and 3i)
edge_slope_scale_max <- df_edge_slopes_system %>% 
  filter(term == "age") %>% 
  pull(t_value) %>% 
  abs %>% 
  max %>% 
  round(1)
```

### A | Normative connectome heatmap through development

```{r Fig5a.function}
f_plot_heatmap <- function(df, pnd) {
  
  # create heatmap plot with bottom system colorbar annotations
  p <- df %>%
    filter(str_detect(subject, "JWD") & timepoint == pnd) %>% 
    group_by(R1, R2, timepoint) %>% 
    summarise(weight = median(weight)) %>% 
    
    mutate(R1 = factor(R1, levels = roi_order),
           R2 = factor(R2, levels = roi_order),
           timepoint = ifelse(timepoint == 300, 230, paste0(timepoint)),
           timepoint = paste0("PND ", timepoint) %>% factor(levels = c("PND 20", "PND 35", "PND 63", "PND 230"))
    ) %>%
    
    # plot
    ggplot(aes(x = R1, y = R2, fill = weight)) +
    geom_tile() +
    
    # add system annotations
    # bottom
    annotate(xmin = c(0.5, system_lines),
             xmax = c(system_lines, length(roi_order) + 0.5),
             ymin = -2.0, ymax = -0.5,
             geom = "rect",
             fill = system_colors) +
    
    # plot aesthetics
    facet_wrap(vars(timepoint), nrow = 1) +
    labs(x = NULL, y = NULL, title = NULL) +
    scale_fill_viridis(na.value = "white", limits = c(0.25, 1.0)) +
    guides(fill = guide_colourbar(title.position = "left", title.hjust = 0.5, title = "w")) +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          legend.position = "none",
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0)
    )
  
    # add system annotations on the left (PND 20)
  if (pnd == 20) {
    p <- p +   
      annotate(ymin = c(0, system_lines),
               ymax = c(system_lines, length(roi_order) + 0.5),
               xmin = -2.0, xmax = -0.5,
               geom = "rect",
               fill = system_colors
      ) #+
      # annotate(geom = "text", x = -2.5, y = system_annotations_noOB,
      #          label = names(system_annotations_noOB), hjust = 1, vjust = 0.25, size = 3
      # ) +
      # coord_cartesian(clip = "off", xlim = c(-1, length(roi_order_noOB)))
  }

  # add legend on the top (PND 300 only)
  if (pnd == 300) {
    p <- p +
      theme(
        legend.position = "top",
        legend.justification = "right",
        legend.key.width = unit(0.5, "cm"),
        legend.key.height = unit(0.125, "cm"),
        legend.margin = margin(t = 0, r = 0, b = -15, l = 0)
      )
  }
  
  # return plot
  return(p)
}
```

```{r Fig5a.plot, fig.width = 4.875, fig.height = 1.65}
## Identify timepoints to plot
timepoints <- df_mind_cortex %>% arrange(timepoint) %>% pull(timepoint) %>% unique

## Generate plots as a list
l_fig5a <- map(timepoints, ~ f_plot_heatmap(df_mind_cortex, pnd = .x))

## Patch together
wrap_plots(l_fig5a, nrow = 1) +
  plot_annotation(title = "A | Normative network at each timepoint") &
  theme(plot.title = element_text(margin = margin(b = -5)))
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "A.normative_connectome_heatmap_by_timepoint", .x), width = 4.875, height = 1.65)
)
```


### B | Normative strength distribution at each timepoint

```{r Fig5b.prep}
## Calculate median degree for each ROI at each timepoint
df_degree_by_timepoint <- df_data_cortex %>% 
  mutate(timepoint = ifelse(timepoint == 300, 230, as.character(timepoint)),
         timepoint = factor(timepoint, levels = c(20, 35, 63, 230))
         #timepoint = paste0("PND ", timepoint) %>% factor(levels = c("PND 20", "PND 35", "PND 63", "PND 230"))
  ) %>% 
  filter(study == "MRC" & feature == "degree") %>% 
  group_by(timepoint, region_of_interest) %>% 
  summarise(degree = median(value, na.rm = TRUE))

## Calculate distribution statistics
kruskal.test(degree ~ timepoint, data = df_degree_by_timepoint)
dunn.test(df_degree_by_timepoint$degree, df_degree_by_timepoint$timepoint, method = "bh")
```

```{r Fig5b.plot, fig.width = 1.625, fig.height = 1.65}
df_degree_by_timepoint %>%
  group_by(timepoint) %>% 
  #mutate(median_degree = median(degree)) %>% 
  
  ggplot(aes(x = timepoint, y = degree)) +
  geom_point(alpha = 0.75, size = 0.3, position = position_jitter(width = 0.15)) +
  geom_boxplot(#aes(color = median_strength, fill = median_strength),
    color = "maroon", fill = "transparent",
    width = 0.5, outlier.shape = NA) +
  #scale_x_discrete(labels = c("PND 20", "PND 35", "PND 63", "PND 230")) +
  #scale_color_viridis() +
  #scale_fill_viridis() +
  labs(x = "PND", y = "Strength (s)",
       title = "B | Strength dist by PND")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "B.degree_distribution_by_timepoint", .x), width = 1.625, height = 1.65)
)
```


### C & D | Brainmaps of nodal slope effect sizes in early and late development

```{r Fig5cd.volumetric.prep}
## Identify non-cortical ROIs
na_rois <- df_whs_atlas %>% 
  anti_join(df_node_slopes_roi) %>% 
  pull(region_of_interest) %>%
  unique

## Set layer order for brain plot based on normative strength
fig5cd_roi_order <- df_node_slopes_roi %>% 
  filter(feature == "degree" & term == "age") %>% 
  group_by(region_of_interest) %>% 
  slice_max(order_by = abs(t_value), n = 1) %>% 
  arrange(abs(t_value)) %>% 
  pull(region_of_interest) %>% 
  unique

## Combine atlas information with ROI weights based on nodal strength slopes
df_fig5cd_volumetric <- df_whs_atlas %>% 
  
  # select plane for plotting
  filter(plane == "sagittal" & hemisphere == "right") %>% 
  dplyr::select(-plane) %>% 
  unnest(cols = c(value)) %>% 
  
  # add degree data
  expand_grid(period = c("early", "aging")) %>% 
  left_join(df_node_slopes_roi %>% 
              filter(feature == "degree" & term == "age") %>% 
              ungroup %>% 
              dplyr::select(region_of_interest, period, t_value, significant)
  ) %>% 
  mutate(period = factor(period, levels = c("early", "aging"))) %>% 
  
  # arrange ROIs for plot
  mutate(region_of_interest = factor(region_of_interest, levels = c(na_rois, fig5cd_roi_order)))
```

```{r Fig5cd.volumetric.plot}
fig5cd_volumetric <- df_fig5cd_volumetric %>% 
  ggplot() +
  geom_polygon(aes(x = x, y = y, group = interaction(region_of_interest, period), fill = t_value), #color = significant), 
               alpha = 1, lwd = 0.25) +
  facet_wrap(vars(period), nrow = 2,
             labeller = as_labeller(c("early" = "C | Early life strength changes Δsdev", "aging" = "D | Age-related strength changes Δsage"))
  ) +
  scale_fill_gradientn(colors = slope_effect_size_scale, limits = c(-degree_slope_scale_max, degree_slope_scale_max), na.value = "lightgray") +
  #scale_color_manual(values = c("1" = "black", "0" = "transparent"), na.value = "transparent", guide = "none") +
  #guides(fill = guide_colorbar(title = "Δk", title.position = "top", title.hjust = 0.5)) +
  #labs(title = "E | Anatomy of developmental degree slopes") +
  theme_void() +
  theme(#plot.title = element_text(face = "bold", size = 12),
        strip.text = element_text(size = 11, face = "bold"),
        strip.clip = "off",
        legend.position = "none",
        # legend.title = element_text(size = 9),
        # legend.text = element_text(size = 8),
        # legend.position = "bottom",
        # legend.justification = "center",
        # legend.key.width = unit(1.5, "cm"),
        # legend.key.height = unit(0.1, "cm"),
        plot.margin = margin(l = 5)
  )
```

```{r fig5cd.flatmap.prep}
## Rotate coordinates by 180 degrees (want the 'front' facing right to align more with anatomic plots)
rotate_180 <- function(geometry) {
  rotation_matrix <- matrix(c(-1, 0, 0, -1), ncol = 2)  # 180-degree counterclockwise rotation
  st_geometry(geometry) <- st_geometry(geometry) * rotation_matrix
  return(geometry)
}
df_flatmap_rot <- rotate_180(df_flatmap)

## Add developmental slopes
df_fig5cd_flatmap <- df_flatmap_rot %>% 
  left_join(df_node_slopes_roi %>% 
              filter(feature == "degree" & term == "age") %>% 
              ungroup %>% 
              dplyr::select(region_of_interest, period, t_value, significant),
            by = join_by(region_of_interest)
  ) %>% 
  filter(!is.na(Swanson_name)) %>% 
  
  # filter for early life slope in left hemisphere, and aging slope in right
  dplyr::filter(
    (period == "early" & hemisphere == "Left") | 
      (period == "aging" & hemisphere == "Right") |
      is.na(region_of_interest)
  )
```

```{r fig5cd.flatmap.plot}
fig5cd_flatmap <- df_fig5cd_flatmap %>% 
  ggplot(mapping = aes(geometry = geometry)) +
  geom_sf(aes(fill = t_value, color = t_value)) +
  scale_fill_gradientn(colors = slope_effect_size_scale, 
                       limits = c(-degree_slope_scale_max, degree_slope_scale_max), 
                       na.value = "white") +
  scale_color_gradientn(colors = slope_effect_size_scale, 
                        limits = c(-degree_slope_scale_max, degree_slope_scale_max), 
                        na.value = "gray",
                        guide = "none") +
  guides(fill = guide_colorbar(title = "Δk", title.position = "top", title.hjust = 0.5)) +
  
  labs(x = NULL, y = NULL, title = NULL) +
  theme(
    legend.position = c(0.5, 1.0),
    legend.direction = "horizontal",
    legend.justification = "center",
    #legend.title = element_text(angle = 90),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(0.2, "cm"),
    legend.margin = margin(l = -20),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
```

```{r Fig5cd.save, fig.width = 4.875, fig.height = 3, eval = FALSE}
layout <- c(
  patchwork::area(t = 1, b = 180, l = 1, r = 90),
  patchwork::area(t = 1, b = 180, l = 91, r = 180)
)
fig5cd_volumetric + 
  fig5cd_flatmap +
  plot_layout(design = layout)
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "CD.strength_slope_brainmaps", .x), width = 4.875, height = 3)
)
```

### E | Correlation of nodal slope effect sizes in early and late development

```{r Fig5e.prep}
## Add system level information
df_fig5e <- df_node_slopes_roi %>% 
  left_join(df_system_hierarchy) %>% 
  filter(feature == "degree" & term == "age") %>% 
  pivot_wider(id_cols = c(system, region_of_interest), names_from = period, values_from = t_value)

## Calculate correation
cor.test(df_fig5e$early, df_fig5e$aging, method = "pearson")
```

```{r Fig5e.plot}
df_fig5e %>% 
  ggplot(aes(x = early, y = aging)) +
  geom_vline(xintercept = 0, lty = 2, color = "darkgrey") +
  geom_hline(yintercept = 0, lty = 2, color = "darkgrey") +
  geom_point(size = 0.5) +
  geom_smooth(method = "lm", color = "maroon", linewidth = 0.5, se = FALSE) +
  stat_cor(color = "maroon", label.sep = "\n", size = 3.5, label.y.npc = "bottom", vjust = 0) +
  labs(x = "Δsdev", y = "Δsage",
       title = "E | Δs correlation"
  ) +
  theme(plot.margin = margin(r = 6))
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "E.strength_slope_correlation", .x), width = small_fig_width, height = small_fig_height)
)
```


### F | Correlation of edge weight slope effect sizes in early and late development

```{r Fig5f.prep}
## Format tibble for plotting
df_fig5f <- df_edge_slopes_system %>% 
  filter(term == "age") %>% 
  pivot_wider(id_cols = system_edge, names_from = period, values_from = t_value)

## Calculate correlation
cor.test(df_fig5f$early, df_fig5f$aging, method = "pearson")
```

```{r Fig5f.plot}
df_fig5f %>% 
  ggplot(aes(x = early, y = aging)) +
  geom_vline(xintercept = 0, lty = 2, color = "darkgrey") +
  geom_hline(yintercept = 0, lty = 2, color = "darkgrey") +
  geom_point(size = 0.5) +
  geom_smooth(method = "lm", color = "maroon", linewidth = 0.5, se = FALSE) +
  stat_cor(color = "maroon", label.sep = "\n", size = 3.5, label.y.npc = "bottom", vjust = 0) +
  labs(x = "Δw_dev", y = "Δw_age",
       title = "F | Δw correlation"
  ) +
  theme(plot.margin = margin(r = 6))
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "F.edge_slope_correlation", .x), width = small_fig_width, height = small_fig_height)
)
```


### G | Circle plots of edge slope effect sizes in early development

```{r Fig5gh.prep}
## Write function to abbreviate names
f_abbreviate_names <- function(names) {
  names %>% 
    str_remove(" region| cortex| formation") %>%
    str_wrap(15) %>% 
    str_replace("frontal", "-\nfrontal") %>%
    str_replace("sensory", "-\nsensory") %>%
    str_replace("splenial", "-\nsplenial") %>%
    str_replace("campal", "-\ncampal")
}

## Abbreviate system names (to save space on plot)
system_order_abrv <- f_abbreviate_names(system_order)
system_colors_abrv <- system_colors
names(system_colors_abrv) <- system_order_abrv
```

```{r Fig5g.prep}
## Create dataframe to generate graph
df_fig5g_graph <- df_edge_slopes_system %>% 
  
  # filter for period & term of interest
  filter(period == "early" & term == "age") %>% 
  ungroup %>% 
  
  # separate edge into constituent systems
  separate(system_edge, into = c("S1", "S2"), sep = " - ") %>% 
  dplyr::select(S1, S2, t_value) %>% 
  
  # edit node names (to save space)
  mutate_at(vars(c(S1, S2)), ~ f_abbreviate_names(.x)) %>% 
  
  # filter for desired effect size
  filter(abs(t_value) > 3.3) %>% 
  
  # order for plotting
  arrange(abs(t_value))

## Create graph
fig5g_graph <- graph_from_data_frame(df_fig5g_graph)

## Order the nodes (for consistency across plots & with the matrix)
fig5g_graph <- permute(fig5g_graph, match(V(fig5g_graph)$name, system_order_abrv))
V(fig5g_graph)$node_label <- names(V(fig5g_graph))
```

```{r Fig5g.plot, fig.height = 4}
ggraph(fig5g_graph, layout = 'linear', circular = TRUE) + 
  geom_edge_arc(aes(color = t_value, width = abs(t_value))) + 
  geom_edge_loop(aes(color = t_value, width = abs(t_value))) + 
  geom_node_text(aes(label = name, angle = node_angle(x, y)), size = 2.5, hjust = "outward") +
  geom_node_point(shape = 21, size = 4, aes(fill = name)) +
  theme_graph() +
  scale_edge_color_gradientn(colors = slope_effect_size_scale, name = "slope", 
                             limits = c(-edge_slope_scale_max, edge_slope_scale_max)
  ) +
  scale_fill_manual(values = system_colors_abrv, guide = "none") +
  scale_edge_width_continuous(range = c(0.1, 2), guide = "none") +
  coord_fixed(xlim = c(-1.75, 1.75), ylim = c(-1.75, 1.75), clip = "off") +
  #labs(title = "H | Early development edge slope") +
  guides(fill = guide_none()) +
  theme_graph() +
  theme(plot.title = element_text(size = 11),
        legend.position = "none")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "G.edge_slope_early_circleplot", .x), width = 3.5, height = 4)
)
```


### H | Circle plots of edge slope effect sizes in late development

```{r Fig5h.prep}
## Create dataframe to generate graph
df_fig5h_graph <- df_edge_slopes_system %>% 
  
  # filter for period & term of interest
  filter(period == "aging" & term == "age") %>% 
  ungroup %>% 
  
  # separate edge into constituent systems
  separate(system_edge, into = c("S1", "S2"), sep = " - ") %>% 
  dplyr::select(S1, S2, t_value) %>% 
  
  # edit node names (to save space)
  mutate_at(vars(c(S1, S2)), ~ f_abbreviate_names(.x)) %>% 
  
  # filter for desired effect size
  filter(abs(t_value) > 3.3) %>% 
  
  # order for plotting
  arrange(abs(t_value))

## Create graph
fig5h_graph <- graph_from_data_frame(df_fig5h_graph)

## Order the nodes (for consistency across plots & with the matrix)
fig5h_graph <- permute(fig5h_graph, match(V(fig5h_graph)$name, system_order_abrv))
V(fig5h_graph)$node_label <- names(V(fig5h_graph))
```

```{r Fig5h.plot, fig.height = 4}
ggraph(fig5h_graph, layout = 'linear', circular = TRUE) + 
  geom_edge_arc(aes(color = t_value, width = abs(t_value))) + 
  geom_edge_loop(aes(color = t_value, width = abs(t_value))) + 
  geom_node_text(aes(label = name, angle = node_angle(x, y)), size = 2.5, hjust = "outward") +
  geom_node_point(shape = 21, size = 4, aes(fill = name)) +
  theme_graph() +
  scale_edge_color_gradientn(colors = slope_effect_size_scale, guide = "none",
                             limits = c(-edge_slope_scale_max, edge_slope_scale_max)
  ) +
  scale_fill_manual(values = system_colors_abrv) +
  scale_edge_width_continuous(range = c(0.1, 2), guide = "none") +
  coord_fixed(xlim = c(-1.75, 1.75), ylim = c(-1.75, 1.75)) +
  #labs(title = "I | Aging edge slope") +
  guides(fill = guide_none(), color = guide_none()) +
  theme(
    plot.title = element_text(size = 11)
  )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "H.edge_slope_aging_circleplot", .x), width = 3.5, height = 4)
)
```


### Generate legend for panels G & H

```{r Fig5fg.legend, fig.width= 4.875, fig.height = 0.5}
circleplot_legend <- (
  df_fig5g_graph %>% 
    ggplot(aes(x = paste0(S1, S2), y = t_value)) +
    geom_point(aes(color = t_value)) +
    guides(colour = guide_colorbar(title = "Δw", title.hjust = 0.5, title.position = "top")) +
    scale_color_gradientn(colors = slope_effect_size_scale, limits = c(-edge_slope_scale_max, edge_slope_scale_max)) +
    theme(legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
          legend.text = element_text(size = 7),
          legend.title = element_text(size = 8),
          legend.direction = "horizontal",
          legend.key.height = unit(0.1, "cm"),
          legend.key.width = unit(2, "cm")
    )
) %>% 
  get_legend()
as.ggplot(circleplot_legend)
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "GH.legend", .x), width = 4.875, height = 0.5)
)
```

